### 2015-11-29 paper/adhoccom.tex
updating the paper given reivew, removing claim on algortihm (there were almost none)

\documentclass[a4paper]{article}

\usepackage{conference}
\usepackage{times}

\usepackage{graphicx}
\usepackage{epstopdf}

\usepackage{caption}
\usepackage{subcaption}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{url}

\usepackage{color}
\newcommand{\todo}[1]{{\small\color{red}[#1]}}

\title{Collaboration in Ad Hoc Teamwork: \\
Ambiguous Tasks, Roles, and Communication\vspace{1em}}

% \author{Anonymous authors for blind review}
\setlength\titlebox{3.8in}

\author{
Jonathan Grizou \\ Flowers Team \\ INRIA - ENSTA ParisTech \\  France \\ jonathan.grizou@inria.fr
\And
Samuel Barrett \\ Kiva Systems \\ North Reading, MA 01864 USA \\ basamuel@kivasystems.com
\AND
Peter Stone \\ Dept. of Computer Science \\ The Univ. of Texas at Austin \\ Austin, TX 78712 USA \\ pstone@cs.utexas.edu
\And
Manuel Lopes \\ Flowers Team \\ INRIA - ENSTA ParisTech \\  France \\ manuel.lopes@inria.fr
}

\newcommand{\env}{E}
\newcommand{\task}{\tau}
\newcommand{\conf}{\kappa}
\newcommand{\com}{\rho}
\newcommand{\gridwidth}{1}
\newcommand{\figwidth}{1}

\DeclareMathOperator*{\argmax}{\arg\!\max}

%CHEAT CODES
% \renewcommand{\baselinestretch}{.98}

\begin{document}

\maketitle

\begin{abstract}
Autonomous agents are becoming increasingly present in our environments, and will increasingly have to collaborate with previously unseen agents. Creating autonomous agents capable of cooperating with previously unfamiliar teammates, known as ``ad hoc teamwork'', has been identified as an important challenge for multiagent systems. Previous research has assumed that either the task, the role of each agent, or the communication protocol among agents is known before the interaction begins. In this paper, we consider these three variables simultaneously and show how an ad hoc agent can fit into a new team while handling ambiguous tasks, roles, and communication protocols. We assume a known distribution of possible tasks, roles, and communication protocols. We present experimental results in the pursuit domain showing that an ad hoc agent can integrate itself into a team by quickly identifying the task, selecting its role on the team, and estimating the communication protocol being used; thus barely impacting the overall team performance compared to a pre-coordinated agent.
\end{abstract}

\input{introduction}

\input{problem}

\input{domain}

\input{method}

\input{results}

\input{conclusion}

\section*{Open Science}

The code developed for this work is available online at \url{https://github.com/jgrizou/adhoc_com}.

\section*{Acknowledgments}

Work partially supported by INRIA, Conseil R\'{e}gional d'Aquitaine, the ERC grant EXPLORERS 24007, and a INRIA Explorer fellowship.

% \clearpage

\bibliographystyle{named}
\bibliography{adhoccom}

\end{document}